quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-10-16 paper-quollr.Rmd
planning implemetation section
Initial writing on adding detour as link plots

@@

The following demonstration of the package's functionality assumes `quollr` has been loaded. To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout. The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.). The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.

To illustrate the workflow, we use the built-in example dataset `scurve`, a $7\text{-}D$ simulated dataset consisting of $1000$ observations. The first three variables define a $3\text{-}D$ S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.

@@

>>>>

To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout. The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.). The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.

@@

The following demonstration of the package's functionality assumes `quollr` has been loaded. To illustrate the workflow, we use the built-in example dataset `scurve`, a $7\text{-}D$ simulated dataset consisting of $1000$ observations. The first three variables define a $3\text{-}D$ S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.

@@

As an alternative to generate tour, the link plots can be generate with `detourr` as well. To achieve this, users can construct the linked visualization manually using the `crosstalk` and `htmltools` packages. The two- and three-panel linked layouts are created by arranging the NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and a dynamic high-dimensional projection view (`langevitour_output` or `detourr_output`) side by side within a flexible grid layout. The `bscols()` function from `crosstalk` synchronizes point selection between these plots, ensuring linked brushing across visual components.

### 2025-10-16 paper-quollr.tex
planning implemetation section
Initial writing on adding detour as link plots

@@

The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout. The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.). The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.

To illustrate the workflow, we use the built-in example dataset \texttt{scurve}, a \(7\text{-}D\) simulated dataset consisting of \(1000\) observations. The first three variables define a \(3\text{-}D\) S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.

@@

>>>>

To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout. The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.). The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.

@@

The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. To illustrate the workflow, we use the built-in example dataset \texttt{scurve}, a \(7\text{-}D\) simulated dataset consisting of \(1000\) observations. The first three variables define a \(3\text{-}D\) S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.

@@

As an alternative to generate tour, the link plots can be generate with \texttt{detourr} as well. To achieve this, users can construct the linked visualization manually using the \texttt{crosstalk} and \texttt{htmltools} packages. The two- and three-panel linked layouts are created by arranging the NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and a dynamic high-dimensional projection view (\texttt{langevitour\_output} or \texttt{detourr\_output}) side by side within a flexible grid layout. The \texttt{bscols()} function from \texttt{crosstalk} synchronizes point selection between these plots, ensuring linked brushing across visual components.
